\mutagen: Coverage-Guided, Property-Based Testing using Exhaustive Structure-Preserving Mutations

### 2022-03-11 papers/fse22/main.tex
NIT
Integrated some of Ale's feedback

\begin{abstract}
Automatically synthesized random data generators are an appealing option when
using property-based testing.
%
There exists a variety of techniques that extract static information from the
codebase to produce random test cases.
%
Sadly, such techniques cannot derive generators capable of producing random
values satisfying complex invariants often needed to test properties with sparse
preconditions.

@@

This work presents \mutagen, a CGPT framework that overcomes such limitations by
generating mutants \emph{exhaustively}.
%
Our tool incorporates heuristics that help to minimize scalability issues as well
as covering the search space in a principled manner.   
%
% schedule mutants based on their
% novelty and to minimize mutating certain subexpressions leading to scalability
% issues of exhaustiveness.
%of large enumeration types more
%than necessary.
%
Our evaluation shows that \mutagen not only outperforms existing CGPT tools but
also finds previously unknown bugs in real-world software.
\end{abstract}

@@

% \vspace{-5pt}

@@

%% PBT and generation of interesting data for free
%
Popularized by \quickcheck \cite{ClaessenH00}, Random Property-Based Testing
(RPBT) is a popular technique for finding bugs
\cite{ClaessenH00,hughes2003erlang, papadakis2011proper, bulwahn2012new,
  denes2014quickchick}.
%
% This is achieved by using randomly generated inputs to instantiate executable
It uses randomly generated inputs to instantiate executable properties encoding
the expected behavior of the system under test.
%
%% Exemplified by frameworks like Haskell's \quickcheck \cite{ClaessenH00} and
%the % Abundant ports of this tool in other programming languages %
%\cite{hughes2003erlang, papadakis2011proper, bulwahn2012new, %
%denes2014quickchick}
%
%% The most common PBT approach is to instantiate the testing properties using %
%randomly generated inputs.
%
% However, users of RPBT in general) are well aware that one of the biggest
% challenges while using RPBT is to provide suitable random data generators
% needed to instantiate the testing properties.
However, a limiting factor of RPBT is the need for the random data generators
used to instantiate the testing properties. 
%
Writing highly-tuned generators capable of exercising every part of a complex
system on a reasonable basis can take several thousand person-hours of trial and
error \cite{lampropoulos2019coverage}.
%
Luckily, there exist several approaches that automatically synthesize random
data generators by reifying the static information present in the codebase,
e.g., data type definitions, application public interfaces (APIs), etc.
\cite{GriecoCB16, DBLP:conf/haskell/MistaRH18, Mista2019GeneratingRS,
DuregardJW12, Lampropoulos2017, Bendkowski2017}.
%
These approaches, however, are unable to synthesize generators capable of
producing data satisfying complex invariants not easily derivable from the
codebase.
%
Randomly generating valid programs to test compilers is an example of this
limitation \cite{chen2020asurvey}, where developers are forced to put
substantial efforts into writing specialized generators by hand \cite{Palka11,
perenyi2020stack, yang2011finding}.
%
Moreover, automatically synthesized generators behave poorly when used to test
properties with sparse preconditions.
%
In this scenario, most tests are simply thrown away before reaching the main
components of the system under test.

@@

In this work, we establish several aspects%
%
\footnote{Some of these limitations were already identified by the authors in
their original work.}
%
of the original CGPT approach by \citeauthor{lampropoulos2019coverage} that
leave considerable room for improvement (see \S \ref{sec:background}).
%
In particular:
%
\begin{inparaenum}
\item automatically derived structure-preserving mutators can turn out
"shallow", ineffective at transforming deep test cases more than superficially;
  % superficial and rely heavily on randomness to produce diverse mutants,
\item the queuing mechanism can cause delays if interesting test cases are
  enqueued frequently and there is no way to prioritize them; and
\item the heuristic used to assign a ``mutation budget'' to each interesting
  test case (often referred to as a \emph{power schedule}) requires fine tuning
  and can be hard to generalize.
%to different kinds of testing properties.
\end{inparaenum}
%
%% Moreover, we replicated the Information-Flow Control (IFC) stack machine case
%% study bundled with \fuzzchick and observed a surprising limitation:
%% %
%% after repeating each experiment 30 times, \fuzzchick could only find 5 out of
%% the 20 injected bugs with a 100\% efficacy.
%% %
%% %% the hardest one being found only around 13\% of the time after an hour of
%% %% testing.
%% %
%% Despite this, \citeauthor{lampropoulos2019coverage} show that \fuzzchick is far
%% better than using normal random testing coupled with na\"ive random generators
%% by comparing the mean-time-to-failure (MTTF) against \emph{QuickChick}.
%% %
%% However, we believe that a more meticulous evaluation ought to take failure to
%% find a counterexample as an important metric when comparing PBT tools, not just
%% mean time to failure --- when one is found.

@@

%% Validation
We validated our ideas using two case studies (described in detail in \S
\ref{sec:casestudies}): the Information-Flow Control (IFC) stack machine used by
\citeauthor{lampropoulos2019coverage} in their original work, and an existing
{Web-Assembly} engine of industrial strength written in Haskell.
%
In both case studies, we compare the effect of the heuristics implemented on top
of the base testing loop of \mutagen.
%
In the IFC stack machine case study, we compare \mutagen against \fuzzchick, the
reference CGPT implementation developed in Coq \cite{lampropoulos2019coverage}.
%
Our results (\S \ref{sec:evaluation}) indicate that \mutagen outperforms
\fuzzchick both in terms of time to failure and failure rate.
%
On the other hand, our WebAssembly case study shows the performance of our tool
in a realistic scenario.
%
There, \mutagen is capable of reliably finding 15 planted bugs in the validator
and interpreter, as well as 3 previously unknown ones.
% bugs that flew under the radar
%of the existing unit test suite of this engine.
%
% This case study also help us to compare the performance of our tool against a
% more traditional RPBT approach that does not rely on code instrumentation.
In this case study we also compare the performance versus the overhead of our
tool (and our custom code instrumentation mechanism) by comparing it against
\quickcheck (black-box).  
%
All in all, our evaluation encompasses more than 600 hours of computing time and
indicates that \emph{testing mutants exhaustively together with our heuristics
to escape scalability issues} can be an appealing technique for finding bugs
reliably without sacrificing speed.%
%
\footnote{A replication package \cite{anon_2021} is available for reviewing
  purposes.}

@@

\begin{verbatim}
prop_insert_inv :: a -> Tree a -> Property
prop_insert_inv x t = balanced t ==> balanced (insert x t)
\end{verbatim}

@@

Readers familiar with RBPT will recognize that using \texttt{genTree} to test
\texttt{prop\_insert\_inv} with \quickcheck directly does not work.
%
In practice, \quickcheck gives up if most of the generated test cases get
discarded.
%
%because they do not satisfy the balanced precondition.
%
Sadly, most of the test cases produced by \texttt{genTree} consist of unbalanced
trees, thus the interesting part of the property (\texttt{balanced (insert x
t)}) is rarely exercised.
%% %
%% In other words, \emph{the precondition of our testing property becomes very
%%   sparse when we use simple, automatically derived generators.}

@@

\citeauthor{lampropoulos2019coverage} demonstrated that CGPT lies in the middle
ground between using random testing with na\"ive automatically derived
generators and complex manually-written ones.
%
%% Their results suggest that CGPT is an appealing technique for finding bugs while
%% still using a mostly automated workflow.
%
However, the authors acknowledge that certain parts of its implementation have
room for improvement. 
%, especially when it comes to the mutator's design.
%
%% Moreover, we replicated the Information-Flow Control (IFC) stack machine case
%
In addition, we observed a notable limitation after replicating the IFC stack
machine case study bundled with \fuzzchick:
%  and observed a notable limitation:
%
\emph{after repeating each experiment 30 times, \fuzzchick could only find 5 out
  of the 20 injected bugs with a 100\% effectivity, the hardest one being found
  only around 13\% of the time after an hour of testing.}
%
Despite this, the authors show that CGPT outperforms random testing coupled with
na\"ive generators by comparing their mean-time-to-failure (MTTF).
%
In turn, we argue that a meticulous evaluation ought to take failure to find a
counterexample as an important metric when comparing PBT tools, not just mean
time to failure --- when one is found.
%
%% In tandem with the lack of efficacy observed when replicating the evaluation of
%% the IFC stack machine case study,

@@

\section{\mutagen: No Mutant Left Behind}
\label{sec:mutagen}

@@

To illustrate this, Fig. \ref{fig:mutagen:mutator} outlines a mutator for the
\texttt{Tree} data type.
%
This definition simply needs to enumerate mutants that transform the outermost
data constructor, hence no recursion is needed here.
%
Moreover, notice how a default value \texttt{def} used to fill the subtrees when
transforming a leaf into a branch.
%
In practice, \texttt{def} corresponds to the simplest expression we can
construct for the mutant to be type-correct.
%
In our example, the default value of a \texttt{Tree} is a leaf containing the
smallest value of the tree payload type, e.g., in the case of trees with
integers \texttt{def} is defined as \texttt{Leaf 0}.
%
%% \texttt{Leaf def}, where
%% the inner \texttt{def} is the default value of the payload.%
(Our implementation uses the type class system to abstract this complexity
away.)
%
Using a small default value, as opposed to a randomly generated one (as done by
the original CGPT), is also inspired by exhaustive bounded testing tools, and
avoids introducing unnecessary randomness when ``growing'' a data constructor
into another.

@@

\vspace{-10pt}
\begin{flalign*}
\mathtt{mut_{s}}(C_i \, x_1 \, x_2 \, \mydots) \!=\! 
  \Vast\{ 
    \mathtt{PURE}(C_j \, x_1^\prime \, x_2^\prime \, \mydots)
    \,\Vast|\,
    \begin{aligned}
      & C_j \!\in\! \{ C_1, C_2, \mydots \} \\
      & x_k^\prime \!\in\! \overline{\mathtt{filter}}(t_k^j, \{x_1, \, x_2, \, \mydots \}) \\  
      & C_j \, x_1^\prime \, x_2^\prime \, \mydots \;\neq\; C_i \, x_1 \, x_2 \, \mydots
    \end{aligned}
  \Vast\} &&
\end{flalign*}
\vspace{-7pt}

@@

\begin{figure}[b]
%% \vspace{-15pt}
\scriptsize
\newcommand\stacked[2]{\underset{\textstyle #2\mathstrut}{#1}}
\begin{align*}
  \nonumber
  %% \qquad\qquad\qquad\qquad
  \left[
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\small \texttt{1}}
      child {node {\small \texttt{2}}
        child {node {\small \texttt{3}}
          child {node {\small \texttt{4}}}
        }
      }
  }
  \right]
  \;
  +
  \;
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\small \texttt{1}}
      child {node {\small \texttt{2}}
        child {node {\small \texttt{3}}
          child {node {\small \texttt{5}}}
        }
      }
  }
  \;
  =
  \;
  \left[
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\color{red} \small \texttt{1}}
      child {node {\color{red} \small \texttt{2}}
        child {node {\color{red} \small \texttt{3}}
          child {node {\small \texttt{4}}}
          child {node {\color{blue} \small \texttt{5}}}
        }
      }
  }
  \right]
  %
  \stacked{{\color{blue} new=1\phantom{x}}}{{\color{red} depth=3}}
  %% \end{equation}
  %% \begin{equation}
  %% \nonumber
  %% \\
  %% \qquad
  \quad
  \left[
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\small \texttt{1}}
      child {node {\small \texttt{2}}
        child {node {\small \texttt{3}}
          child {node {\small \texttt{4}}}
          child {node {\small \texttt{5}}}
        }
      }
  }
  \right]
  \;
  +
  \;
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\small \texttt{1}}
      child {node {\small \texttt{2}}
        child {node {\small \texttt{6}}
          child {node {\small \texttt{7}}
          }
        }
      }
  }
  \;
  =
  \;
  \left[
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.75cm}
  ]{
    \node {\color{red} \small \texttt{1}}
      child {node {\color{red} \small \texttt{2}}
        child {node {\small \texttt{3}}
          child {node {\small \texttt{4}}}
          child {node {\small \texttt{5}}}
        }
        child {node {\color{blue}\small \texttt{6}}
          child {node {\color{blue}\small \texttt{7}}}
        }
      }
  }
  \right]
  %
  \stacked{{\color{blue} new=2\phantom{x}}}{{\color{red} depth=2}}
\end{align*}
\vspace{-10pt}
\caption{\label{fig:mutagen:tracelog}Inserting two new execution traces into the
  internal execution trace log (represented using brackets).}
\end{figure}

@@

The second case study evaluates \mutagen in a realistic scenario, and targets
\textit{haskell-wasm} \cite{haskellwasm}, an existing WebAssembly engine of
industrial strength.
%
%
Unfortunately, comparing \mutagen with \fuzzchick on this case study has been
out of the scope of this work, since porting \textit{haskell-wasm} into Coq, and
later adapting our test suite to work with \fuzzchick requires a subtantial
effort.
%
Instead, we compare \mutagen against \quickcheck, evaluating its effectiveness
versus the relative overhead of the code instrumentation used by our tool.

@@

This abstract stack machine consists of four main elements: a program counter, a
stack, and data- and instruction memories.
%
Moreover, every runtime value is labeled with a security level, i.e., L (for
``low'' or public) or H (for ``high'' or secret).
%
Labels form a trivial 2-lattice where information can either stay at the same
level, or public information ``can flow'' to secret one but not the opposite
(represented using the familiar $\sqsubseteq$ binary operator).
%
Security labels are propagated throughout the execution of the program whenever
the machine executes an instruction.
%
These instructions are defined as:
%
\begin{verbatim}
data Instr = Nop | Push Int | Call Int | Ret | Add | Load | Store | Halt
\end{verbatim}

@@

%% \vspace{-10pt}
\begin{center}
\scriptsize
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Instruction} & \textbf{Precondition Check} & \textbf{Final PC Label} & \textbf{Final Result Label} \\
\hline
\texttt{Store} & $l_{pc} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
\texttt{Store} & $l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
\hline
\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $\bot$ & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
\hline
\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{pc} \vee l_{p}$ \\
\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{v'} \vee l_{p}$ \\
\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{v'} \vee l_{pc}$ \\
\hline
\end{tabular}
\end{center}

@@

\begin{table}[b]
\scriptsize
\begin{tabular}{|c|c|c|l|}
\hline
\textbf{Id}
& \textbf{Subsystem}
& \textbf{Category}
& \textbf{Description} \\
\hline
1
& Validator
& Bug
& Invalid memory alignment validation \\
\hline
2
& Validator
& Discrepancy
& Validator accepts blocks returning multiple values  \\
\hline
3
& Interpreter
& Bug
& Instance function invoker silently ignores arity mismatch \\
\hline
4
& Interpreter
& Bug
& Allowed out-of-bounds memory access \\
\hline
5
& Interpreter
& Discrepancy
& NaN reinterpretation does not follow reference \\
\hline
\end{tabular}
\caption{\label{table:wasm:bugs}Issues found by \mutagen in
  \textit{haskell-wasm}.}
\vspace{-10pt}
\end{table}

>>>>

\begin{abstract}
Automatically synthesized random data generators are an appealing option when
using property-based testing.
%
There exists a variety of techniques that extract static information from the
codebase to produce random test cases.
%
% Sadly, such techniques cannot derive generators capable of producing random
%
Unfortunately, such techniques suffer from severe limitations when used to
produce random test cases satisfying complex invariants, which are often needed
to test properties with sparse preconditions.

@@

This work presents \mutagen, a CGPT framework that overcomes such limitations by
generating mutants \emph{exhaustively}.
%
Our tool incorporates heuristics that help to minimize scalability issues as well
as cover the search space in a principled manner.   
%
% schedule mutants based on their
% novelty and to minimize mutating certain subexpressions leading to scalability
% issues of exhaustiveness.
%of large enumeration types more
%than necessary.
%
Our evaluation shows that \mutagen not only outperforms existing CGPT tools but
also finds previously unknown bugs in real-world software.
\end{abstract}

@@

% \vspace{-10pt}

@@

%% PBT and generation of interesting data for free
%
% Popularized by \quickcheck \cite{ClaessenH00},
%
Random Property-Based Testing (RPBT) is a popular technique for finding bugs
\cite{ClaessenH00,hughes2003erlang, papadakis2011proper, bulwahn2012new,
denes2014quickchick}.
%
% This is achieved by using randomly generated inputs to instantiate executable
It uses randomly generated inputs to instantiate executable properties encoding
the expected behavior of the system under test.
%
%% Exemplified by frameworks like Haskell's \quickcheck \cite{ClaessenH00} and
%the % Abundant ports of this tool in other programming languages %
%\cite{hughes2003erlang, papadakis2011proper, bulwahn2012new, %
%denes2014quickchick}
%
%% The most common PBT approach is to instantiate the testing properties using %
%randomly generated inputs.
%
% However, users of RPBT in general) are well aware that one of the biggest
% challenges while using RPBT is to provide suitable random data generators
% needed to instantiate the testing properties.
However, a limiting factor of RPBT is the need for the random data generators
used to instantiate the testing properties. 
%
Writing highly-tuned generators capable of exercising every part of a complex
system on a reasonable basis can take several thousand person-hours of trial and
error \cite{lampropoulos2019coverage}.
%
Luckily, there exist several approaches that automatically synthesize random
data generators by reifying the static information present in the codebase,
e.g., data type definitions and application public interfaces (APIs).
\cite{GriecoCB16, DBLP:conf/haskell/MistaRH18, Mista2019GeneratingRS,
DuregardJW12, Lampropoulos2017, Bendkowski2017}.
%
These approaches, however, are unable to synthesize generators capable of
producing data satisfying complex invariants not easily derivable from the
codebase.
%
Randomly generating valid programs to test compilers is an example of this
limitation \cite{chen2020asurvey}, where developers are forced to put
substantial efforts into writing specialized generators by hand \cite{Palka11,
perenyi2020stack, yang2011finding}.
%
Moreover, automatically synthesized generators behave poorly when used to test
properties with sparse preconditions.
%
In this scenario, most tests are simply thrown away before reaching the main
components of the system under test.

@@

In this work, we establish several aspects%
%
% \footnote{Some of these limitations were already identified by the authors in
% their original work.}
%
of the original CGPT approach by \citeauthor{lampropoulos2019coverage} that
leave considerable room for improvement (see \S \ref{sec:background}).
%
In particular:
%
\begin{inparaenum}
\item automatically derived structure-preserving mutators can turn out
"shallow", ineffective at transforming deep test cases more than superficially;
  % superficial and rely heavily on randomness to produce diverse mutants,
\item the queuing mechanism can cause delays if interesting test cases are
  enqueued frequently and there is no way to prioritize them; and
\item the heuristic used to assign a ``mutation budget'' to each interesting
  test case (often referred to as a \emph{power schedule}) requires fine tuning
  and can be hard to generalize.
%to different kinds of testing properties.
\end{inparaenum}

%% Moreover, we replicated the Information-Flow Control (IFC) stack machine case
%% study bundled with \fuzzchick and observed a surprising limitation:
%% %
%% after repeating each experiment 30 times, \fuzzchick could only find 5 out of
%% the 20 injected bugs with a 100\% efficacy.
%% %
%% %% the hardest one being found only around 13\% of the time after an hour of
%% %% testing.
%% %
%% Despite this, \citeauthor{lampropoulos2019coverage} show that \fuzzchick is far
%% better than using normal random testing coupled with na\"ive random generators
%% by comparing the mean-time-to-failure (MTTF) against \emph{QuickChick}.
%% %
%% However, we believe that a more meticulous evaluation ought to take failure to
%% find a counterexample as an important metric when comparing PBT tools, not just
%% mean time to failure --- when one is found.

@@

%% Validation
We validated our ideas using two case studies which require generating
highly-structured inputs (described in detail in \S \ref{sec:casestudies}): the
Information-Flow Control (IFC) stack machine used by
\citeauthor{lampropoulos2019coverage} in their original work, and an existing
{Web-Assembly} engine of industrial strength written in the Haskell programming
language.
%
In both case studies, we compare the effect of the heuristics implemented on top
of the base testing loop of \mutagen.
%
In the IFC stack machine case study, we compare \mutagen against \fuzzchick, the
reference CGPT implementation developed for the Coq proof assistant
\cite{lampropoulos2019coverage}.
%
Our results (\S \ref{sec:evaluation}) indicate that \mutagen outperforms
\fuzzchick both in terms of time to failure and failure rate.
%
On the other hand, our WebAssembly case study shows the performance of our tool
in a realistic scenario.
%
There, \mutagen is capable of reliably finding 15 planted bugs in the validator
and interpreter, as well as 3 previously unknown ones.
% bugs that flew under the radar
%of the existing unit test suite of this engine.
%
% This case study also help us to compare the performance of our tool against a
% more traditional RPBT approach that does not rely on code instrumentation.
In this case study we also evaluate the performance versus the overhead of our
tool (and our custom code instrumentation mechanism) by comparing it against
\quickcheck \cite{ClaessenH00}, the most widely used RPBT tool in Haskell.  
%
All in all, our evaluation encompasses more than 600 hours of computing time and
indicates that \emph{testing mutants exhaustively together with our heuristics
to escape scalability issues} can be an appealing technique for finding bugs
reliably without sacrificing speed.%
%
\footnote{A replication package \cite{anon_2021} is available for reviewing
  purposes.}

@@

\begin{verbatim}
prop_insert :: a -> Tree a -> Property
prop_insert x t = balanced t ==> balanced (insert x t)
\end{verbatim}

@@

Readers familiar with RBPT will recognize that using \texttt{genTree} to
generate test cases to test \texttt{prop\_insert} with \quickcheck directly does
not work.
%
In practice, \quickcheck gives up if most of the generated test cases get
discarded.
%
%because they do not satisfy the balanced precondition.
%
Sadly, most of the test cases produced by \texttt{genTree} consist of unbalanced
trees, thus the interesting part of the property (\texttt{balanced (insert x
t)}) is rarely exercised.
%% %
%% In other words, \emph{the precondition of our testing property becomes very
%%   sparse when we use simple, automatically derived generators.}

@@

\citeauthor{lampropoulos2019coverage} demonstrated that CGPT lies in the middle
ground between using random testing with na\"ive automatically derived
generators and complex manually-written ones.
%
%% Their results suggest that CGPT is an appealing technique for finding bugs while
%% still using a mostly automated workflow.
%
% However, the authors acknowledge that certain parts of its implementation have
% room for improvement. 
%, especially when it comes to the mutator's design.
%
%% Moreover, we replicated the Information-Flow Control (IFC) stack machine case
%
% In addition, 
However, we observed a notable limitation after replicating the IFC stack
machine case study bundled with \fuzzchick:
%  and observed a notable limitation:
%
\emph{after repeating each experiment 30 times, \fuzzchick could only find 5 out
  of the 20 injected bugs with a 100\% effectivity, the hardest one being found
  only around 13\% of the time after an hour of testing.}
%
Despite this, the authors show that CGPT outperforms random testing coupled with
na\"ive generators by comparing their mean-time-to-failure (MTTF).
%
In turn, we argue that a meticulous evaluation ought to take failure to find a
counterexample as an important metric when comparing PBT tools, not just mean
time to failure.
% --- when one is found.
%
%% In tandem with the lack of efficacy observed when replicating the evaluation of
%% the IFC stack machine case study,

@@

\section{\mutagen}
\label{sec:mutagen}

@@

To illustrate this, Fig. \ref{fig:mutagen:mutator} outlines a mutator for the
\texttt{Tree} data type.
%
This definition simply needs to enumerate mutants that transform the outermost
data constructor, hence no recursion is needed here.
%
Moreover, notice how a default value \texttt{def} used to fill the subtrees when
transforming a leaf into a branch.
%
In practice, \texttt{def} corresponds to the simplest expression we can
construct for the mutant to be type-correct.
%
In our example, the default value of a \texttt{Tree} is a leaf containing the
smallest value of the tree payload type, e.g., in the case of trees with
integers \texttt{def} is defined as \texttt{Leaf 0}.
%
%% \texttt{Leaf def}, where
%% the inner \texttt{def} is the default value of the payload.%
% (Our implementation uses the type class system to abstract this complexity
% away.)
%
Using a small default value, as opposed to a randomly generated one (as done by
the original CGPT), is also inspired by exhaustive bounded testing tools, and
avoids introducing unnecessary randomness when ``growing'' a data constructor
into another.

@@

\vspace{-10pt}
\begin{flalign*}
\mathtt{mut_{s}}(C_i \, x_1 \, x_2 \, \mydots) \!=\! 
  \Vast\{ 
    \mathtt{PURE}(C_j \, x_1^\prime \, x_2^\prime \, \mydots)
    \Vast|\,
    \begin{aligned}
      & C_j \!\in\! \{ C_1, C_2, \mydots \} \\
      & x_k^\prime \!\in\! \overline{\mathtt{filter}}(t_k^j, \{x_1, \, x_2, \, \mydots \}) \\  
      & C_j \, x_1^\prime \, x_2^\prime \, \mydots \;\neq\; C_i \, x_1 \, x_2 \, \mydots
    \end{aligned}
  \!\Vast\} &&
\end{flalign*}
\vspace{-7pt}

@@

\begin{figure}[b]
\vspace{-5pt}
\scriptsize
\newcommand\stacked[2]{\underset{\textstyle #2\mathstrut}{#1}}
\begin{align*}
  \nonumber
  %% \qquad\qquad\qquad\qquad
  \left[
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\small \texttt{1}}
      child {node {\small \texttt{2}}
        child {node {\small \texttt{3}}
          child {node {\small \texttt{4}}}
        }
      }
  }
  \right]
  \;
  +
  \;
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\small \texttt{1}}
      child {node {\small \texttt{2}}
        child {node {\small \texttt{3}}
          child {node {\small \texttt{5}}}
        }
      }
  }
  \;
  =
  \;
  \left[
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\color{red} \small \texttt{1}}
      child {node {\color{red} \small \texttt{2}}
        child {node {\color{red} \small \texttt{3}}
          child {node {\small \texttt{4}}}
          child {node {\color{blue} \small \texttt{5}}}
        }
      }
  }
  \right]
  %
  \stacked{{\color{blue} new=1\phantom{x}}}{{\color{red} depth=3}}
  %% \end{equation}
  %% \begin{equation}
  %% \nonumber
  %% \\
  %% \qquad
  \quad
  \left[
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\small \texttt{1}}
      child {node {\small \texttt{2}}
        child {node {\small \texttt{3}}
          child {node {\small \texttt{4}}}
          child {node {\small \texttt{5}}}
        }
      }
  }
  \right]
  \;
  +
  \;
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\small \texttt{1}}
      child {node {\small \texttt{2}}
        child {node {\small \texttt{6}}
          child {node {\small \texttt{7}}
          }
        }
      }
  }
  \;
  =
  \;
  \left[
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.75cm}
  ]{
    \node {\color{red} \small \texttt{1}}
      child {node {\color{red} \small \texttt{2}}
        child {node {\small \texttt{3}}
          child {node {\small \texttt{4}}}
          child {node {\small \texttt{5}}}
        }
        child {node {\color{blue}\small \texttt{6}}
          child {node {\color{blue}\small \texttt{7}}}
        }
      }
  }
  \right]
  %
  \stacked{{\color{blue} new=2\phantom{x}}}{{\color{red} depth=2}}
\end{align*}
\vspace{-15pt}
\caption{\label{fig:mutagen:tracelog}Inserting two new execution traces into the
  internal execution trace log (represented using brackets).}
\end{figure}

@@

The second case study evaluates \mutagen in a realistic scenario, and targets
\textit{haskell-wasm} \cite{haskellwasm}, an existing WebAssembly engine of
industrial strength.
%
%
Unfortunately, comparing \mutagen with \fuzzchick on this case study has been
out of the scope of this work, since porting \textit{haskell-wasm} into Coq, and
later adapting our test suite to work with \fuzzchick requires a subtantial
effort.
%
Instead, we compare \mutagen against \quickcheck, evaluating its effectiveness
versus the relative overhead of our custom code instrumentation.

@@

This abstract stack machine consists of four main elements: a program counter, a
stack, and data- and instruction memories.
%
Moreover, every runtime value is labeled with a security level, i.e., L (for
``low'' or public) or H (for ``high'' or secret).
%
Labels form a trivial 2-lattice where information can either stay at the same
level, or public information ``can flow'' to secret one but not the opposite
(represented using the familiar $\sqsubseteq$ binary operator).
%
Security labels are propagated throughout the execution of the program whenever
the machine executes an instruction.
%
These instructions are defined as:

\vspace{5pt}
\begin{verbatim}
data Instr = Nop | Push Int | Call Int | Ret | Add | Load | Store | Halt
\end{verbatim}

@@

%% \vspace{-10pt}
\begin{center}
\scriptsize
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Instruction} & \textbf{Precondition Check} & \textbf{Final PC Label} & \textbf{Final Result Label} \\
\hline
\texttt{Store} & \color{red}{$l_{pc} \sqsubseteq l_{v}$} & $l_{pc}$ & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
\texttt{Store} & \color{red}{$l_{p} \sqsubseteq l_{v}$} & $l_{pc}$ & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
\hline
\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & \color{red}{$\bot$} & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
\hline
\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & \color{red}{$l_{pc} \vee l_{p}$} \\
\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & \color{red}{$l_{v'} \vee l_{p}$} \\
\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & \color{red}{$l_{v'} \vee l_{pc}$} \\
\hline
\end{tabular}
\end{center}

@@

\begin{table}[b]
% \vspace{10pt}
\scriptsize
\begin{tabular}{|c|c|c|l|}
\hline
\textbf{Id}
& \textbf{Subsystem}
& \textbf{Category}
& \textbf{Description} \\
\hline
1
& Validator
& Bug
& Invalid memory alignment validation \\
\hline
2
& Validator
& Discrepancy
& Validator accepts blocks returning multiple values  \\
\hline
3
& Interpreter
& Bug
& Instance function invoker silently ignores arity mismatch \\
\hline
4
& Interpreter
& Bug
& Allowed out-of-bounds memory access \\
\hline
5
& Interpreter
& Discrepancy
& NaN reinterpretation does not follow reference \\
\hline
\end{tabular}
\caption{\label{table:wasm:bugs}Issues found by \mutagen in
  \textit{haskell-wasm}.}
% \vspace{-10pt}
\end{table}

\begin{table}[b]
\scriptsize
\begin{tabular}{|c|c|l|}
\hline
\textbf{Id}
& \textbf{Subsystem}
& \textbf{Description} \\
\hline
1
& Validator
& Wrong if-then-else type validation on else branch \\
\hline
2
& Validator
& Wrong stack type validation \\
\hline
3
& Validator
& Removed function type mismatch assertion \\
\hline
4
& Validator
& Removed max memory instances assertion \\
\hline
5
& Validator
& Removed function index out-of-range assertion \\
\hline
6
& Validator
& Wrong type validation on \texttt{i64.eqz} instruction \\
\hline
7
& Validator
& Wrong type validation on \texttt{i32} binary operations \\
\hline
8
& Validator
& Removed memory index out-of-range assertion \\
\hline
9
& Validator
& Wrong type validation on \texttt{i64} constants \\
\hline
10
& Validator
& Removed alignment validation on \texttt{i32.load} instruction \\
\hline
11
& Interpreter
& Wrong interpretation of \texttt{i32.sub} instruction \\
\hline
12
& Interpreter
& Wrong interpretation of \texttt{i32.lt\_u} instruction \\
\hline
13
& Interpreter
& Wrong interpretation of \texttt{i32.shr\_u} instruction \\
\hline
14
& Interpreter
& Wrong local variable initialization \\
\hline
15
& Interpreter
& Wrong memory address casting on \texttt{i32.load8\_s} instruction \\
\hline
\end{tabular}
\caption{\label{table:wasm:injectedbugs}Bugs injected into
  \textit{haskell-wasm}.}
\vspace{-20pt}
\end{table}
